Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
in a mess

\begin{proof}
  That $\graphof A$ is a logical cograph follows immediately from the definition and Theorem~\ref{thm:cograph}.
  The fact that every binder of $\graphof A$ is legal can be proved by
  structural induction on $A$.
  %% :
  %% \begin{itemize}
  %%   \item Trivial if $A$ is an atom.
  %%   \item The induction step uses the following fact:
  %%         if $M$ is a module of $\gG$ and $M' \subseteq M$, then $M'$ is a
  %%         module of $\gG$ if and only if it is a module of $\gG[X]$.
  %% \end{itemize}
  %% \todo{}\lutz{maybe mention also fn 2-5 in Dominic's arxiv-paper}
\end{proof}
%%\juihsuan{I'm still looking for a simple proof}
%
\begin{remark}
Note that $\graphof A$ is not necessarily a fograph if $A$ is not
rectified. If $A = (\forall x. px) \vlor (\forall x. qx)$, then
$\graphof A = \single x \ \single px \ \single x \ \single qx$,
the scope of each $x$-binder contains all the vertices, in particular,
the other $x$-binder. On the other hand, there are non-rectified
formulas which are translated to fographs by $\graphof\cdot$. For
example, in the graph of $(\exists x. px) \vlor (\exists x. qx)$,
both $x$-binders are legal, as they are not in each other's
scope:
$\edgepairpic{x}{px}\mkern9mu\edgepairpic{x}{qx}$.
\end{remark}
%
%% For this reason, we call a formula
%% \bfit{clean} if it does not contain subformulas of the form $(\forall
%% x.A)\vlor(\forall x.B)$ or $(\exists x.A)\vlan(\exists x.B)$,
%% \todo{need the four cases as in clear definition in fopws? Consider
%% $(\forall x.A)\vlor(B\vlor (\forall x.C))$. But then Lemma 6 doesn't hold.
%% (fopws proves only that there's a surjection from clear formulas onto fographs,
%% whose kernel is the six formula equalities)} and no
%% $x$-quantified formula occurs as subformula of another $x$-quantified
%% formula. Then we have:
%
%% \begin{lemma}
%%   If $A$ is clean iff\/ $\graphof A$ is a fograph.
%% \end{lemma}
%
%% \begin{proof}
%%   Induction on $A$, using  Theorem~\ref{thm:cograph}.
%% \end{proof}
%
%% Even though for every formula $A$ we can obtain an
%% equivalent clean formula by simply renaming some bound variables, this
%% is not unique up to $\alpha$-conversion, as is the case for
%% rectified formulas.
%
\noindent We define a congruence relation $\fequ$ on formulas by the following equations:
\begin{equation}
  \label{eq:fequ}
  \begin{array}{c@{\hskip1.5em}c}
    A \vlan B \fequ B \vlan A
    &
    (A \vlan B) \vlan C \fequ A \vlan (B \vlan C) 
    \\
    A \vlor B \fequ B \vlor A
    &
    (A \vlor B) \vlor C \fequ A \vlor (B \vlor C)
    \\
    \forall x . \forall y . A \fequ \forall y . \forall x . A
    &
    \forall x . (A \vlor B) \fequ (\forall x . A) \vlor B
    \\
    \exists x . \exists y . A \fequ \exists y . \exists x . A
    &
    \exists x . (A \vlan B) \fequ (\exists x . A) \vlan B
  \end{array}
\end{equation}
where $x$ must not be free in $B$ in the last two equations.
Two formulas $A$ and $B$ are \bfit{equivalent} if $A \fequ B$. The following
theorem shows that the set of fographs can been seen as the quotient
$\FORM/\mathord\fequ$.
\begin{thm}
  Let $A,B$ be rectified formulas. Then
  $$
  A\fequ B \iff \graphof A =\graphof B
  $$
\end{thm}

@@

Let $\Pi$ be a $\FOMLL$ proof of a rectified sequent $\sqn\Gamma$ not containing $\fff$. We now show
how $\Pi$ is translated into a linked fograph
$\fographof\Pi=\tuple{\graphof\Gamma,\linkingof\Pi}$. We proceed
inductively, making a case analysis on the last rule in $\Pi$. At the
same time we are constructing a dualizer $\dsubstof\Pi$, so that in the
end we can conclude that $\fographof\Pi$ is in fact a fonet.
\begin{enumerate}
  \addtolength\itemsep{.7ex}
\item $\Pi$ is $\upsmash{\vlinf{\axr}{}{\sqn{a,\cneg a}}{}}$ : Then the only
  link is $\set{a, \dual{a}}$, and $\dsubstof\Pi$ is empty.
\item $\Pi$ is $\smash{\vlinf{\ttt}{}{\sqn{\ttt}}{}}$ : Then
  $\linkingof\Pi$ and $\dsubstof\Pi$ are both empty.
\item The last rule in $\Pi$ is $\vliinf{\mixr}{}{\sqn{\Gamma',\Gamma''}}{\sqn\Gamma'}{\sqn\Gamma''}$ :
  By induction hypothesis, we have proofs $\Pi'$ and $\Pi''$ of $\Gamma'$ and $\Gamma''$, respectively. We have
  $\graphof{\Gamma}=\graphof{\Gamma'}+\graphof{\Gamma''}$ and we can let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}\cup\linkingof{\Pi''}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}\cup\dsubstof{\Pi''}$.
\item The last rule in $\Pi$ is $\vlinf{\vlor}{}{\sqn{\Gamma_1,A\vlor
    B}}{\sqn{\Gamma_1,A,B}}$ : By induction hypothesis, there is
  a proof $\Pi'$ of $\Gamma'={\Gamma_1,A,B}$. We have
  $\graphof{\Gamma}=\graphof{\Gamma'}$ and let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}$.
\item The last rule in $\Pi$ is $\vliinf{\vlan}{}{\sqn{\Gamma_1,A\vlan
    B, \Gamma_2}}{\sqn{\Gamma_1,A}}{\sqn{B,\Gamma_2}}$ : By induction
  hypothesis, we have proofs $\Pi'$ and $\Pi''$ of
  $\Gamma'=\Gamma_1,A$ and $\Gamma''=B,\Gamma_2$, respectively. We
  have $\graphof{\Gamma}=\graphof{\Gamma_1}+\mbox{$(\graphof A\times\graphof
  B)$}+\graphof{\Gamma_2}$ and we let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}\cup\linkingof{\Pi''}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}\cup\dsubstof{\Pi''}$.
\item The last rule in $\Pi$ is
  $\vlinf{\exists}{}{\sqn{\Gamma_1,\exists x.A}}
  {\sqn{\Gamma_1,A\ssubst{x}{t}}}$ : By induction hypothesis, there is
  a proof $\Pi'$ of $\Gamma'={\Gamma_1,A\ssubst{x}{t}}$.  For each atom in
  $\Gamma'=\Gamma_1, A \ssubst{x}{t}$, there is a corresponding atom
  in $\Gamma=\Gamma_1, \exists x.A$. We can therefore define the
  linking $\linkingof{\Pi}$ from the linking $\linkingof{\Pi'}$ via
  this correspondence. Then, we let $\dsubstof\Pi$ be
  $\dsubstof{\Pi'}+\ssubst{x}{t}$. Since $\Gamma$ is rectified $x$
  does not yet occur in $\dsubstof{\Pi'}$. Hence $\dsubstof\Pi$ is a
  dualizer of $\fographof\Pi$.
\item The last rule in $\Pi$ is $\vlinf{\forall}{\text{($x$ not free
    in $\Gamma_1$)}}{\sqn{\Gamma_1,\forall x.A}}{\sqn{\Gamma_1,A}}$ :
  By induction hypothesis, there is a proof $\Pi'$ of
  $\Gamma'={\Gamma_1,A}$, which has the same atoms as in
  $\Gamma=\Gamma_1, \forall x.A$.  Hence, we can let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}$. 
\end{enumerate}
%\end{itemize}

@@

\begin{lemma}\label{lem:skew->cw}
  Let $\gA$ and $\gB$ be fographs, let $\phi\colon\gA\to\gB$ be a skew
  bifibration, and let $A$ and $B$ be formulas with $\graphof A=\gA$
  and $\graphof B=\gB$. Then there are derivations
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wrD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{B}{
        \vlhy{A}}}
    \quand
    \vlderivation{
      \vlde{\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
        \vlhy{A\rsubstof\phi}}}
  \end{equation*}
  such that $\mapof{\rectif\Deri}=\phi$ and $\rectif{\Deri}$ is a
  rectification of $\Deri$, and $\rsubstof\phi$ is the substitution
  induced by $\phi$.
\end{lemma}

@@

\begin{proof}[Proof of Lemma~\ref{lem:skew->cw}]
  By Lemma~\ref{lem:PE} we have $\gA=\graphof{\PE A}$ and
  $\gB=\graphof{\PE B}$.  Let $\vB'\subseteq\vB$ be the image of
  $\phi$, and let $\gB_1$ be the subgraph of $\gB$ induced by
  $\vB'$. Hence, we have two maps $\phi''\colon\gA\to\gB_1$ being a
  surjection and $\phi'\colon \gB_1\to\gB$ being an injection that
  reflects edges.  Both, $\phi'$ and $\phi''$ remain skew
  bifibrations. Let us first look at $\phi'$.  Let $\PEp B_1$ be the
  propositional formula obtained from $\PE B$ by removing all atoms
  that are not represented by vertices in $\vB'$. Then $\graphof{\PEp
    B_1}=\gB_1$.  By \cite[Proposition~7.6.1]{str:07:RTA}, we have a
  derivation $\vlder{\set{\wrD,\PE\fequ}}{\PE\Deri_1}{\PE B}{\PEp B_1}$.
  A subformula of $\PE B$ is called \emph{weak} if it has been
  introduced by a weakening. All subformulas of a weak formula are
  also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  \emph{weak pair} if $\PE B\fequ\Scons{B'\vlor B''}$ for some context
  $\Sconhole$. We can assume without loss of generality that
  whenever weak subformulas $B'$ and $B''$ form a weak pair, they have
  been introduced by the same instance of $\wrD$ in
  $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  brought into this form by simple rule permutations, and then
  collapsing several weakenings into a single one.}  Now we show that
  $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  delete all atoms in the scope of the corresponding quantifier,
  because $\phi'$ is a fibration on the binding graph. Hence, each line
  in $\PE\Deri_1$ is the propositional encoding $\PE P$ of a
  first-order formula $P$.  We now have to show
  that each instance of $\wrD$ is indeed a correct
  application in first-order logic. For this we have to inspect the
  cases a weakening happens inside a subformula $x\vlor C$ or $x\vlan
  C$ in $\PE\Deri_1$. There are the following cases:
  \begin{equation*}%\small
    \vlinf{}{}{S\cons{x\vlor D\vlor C}}{S\cons{x\vlor C}}
    \qquad
    \vlinf{}{}{S\cons{x\vlan (D\vlor C)}}{S\cons{x\vlan C}}
    \qquad
    \vlinf{}{}{S\cons{(x\vlor D)\vlan C}}{S\cons{x\vlan C}}
  \end{equation*}
  In the first case the weakening happens inside the scope of a
  $\forall$-quantifier, and in the second case inside the scope of a
  $\exists$-quantifier. Both cases are unproblematic in first-order
  logic. However, in the third case an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur.
  Thus we have a first order derivation
  $\vlder{\set{\wrD,\fequ}}{\Deri_1}{B}{B_1}$ with $\PE B_1=\PEp B_1$.

@@

\begin{equation*}
\vlderivation{
  \vliin{\vlan}{}{\sqns{\Gamma, A \vlan B, \Delta}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A}}{
      \vlhy{\sqns{\Gamma}}}}{
    \vlhy{\sqns{B, \Delta}}}}
\leadsto
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, A \vlan B, \Delta}}{
    \vlin{\weakr}{}{\vlvdots}{
      \vlhy{\sqns{\Gamma}}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\vlor}{}{\sqns{\Gamma, A \vlor B}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A, B}}{
      \vlhy{\sqns{\Gamma, A}}}}}
\leadsto
\vlderivation{
  \vlin{\wrD}{}{\sqns{\Gamma, A \vlor B}}{
    \vlhy{\sqns{\Gamma, A}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\exists}{}{\sqns{\Gamma, \exists x.A}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A\ssubst{x}{t}}}{
      \vlhy{\sqns{\Gamma}}}}}
\leadsto
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, \exists x.A}}{
    \vlhy{\sqns{\Gamma}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\forall}{}{\sqns{\Gamma, \forall x.A}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A}}{
      \vlhy{\sqns{\Gamma}}}}}
\leadsto
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, \forall x.A}}{
    \vlhy{\sqns{\Gamma}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\conr}{}{\sqns{\Gamma, A}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A, A}}{
      \vlhy{\sqns{\Gamma, A}}}}}
\leadsto
\vlderivation{
  \vlhy{\sqns{\Gamma, A}}}
\end{equation*}
Note that in the case of $\vlor$, we use the deep rule $\wrD$ which can be
permuted under all the rules. By using these rewriting rules, we can eventually get
a derivation with all the instances of $\weakr$ and $\wrD$ at the bottom. Now observe
that the instances of $\conr$ in $\Deri$ can be transformed using the following
rule:

@@

Knowing that $\cD$ can be permuted under all the rules of $\FOMLL$, we
eventually obtain a derivation:
\begin{equation*}
\vlderivation{
  \vlde{\set{\weakr, \wrD,\cD,\fequ}}{\Deri'_2}{\sqns{\Gamma}}{
    \vltrl{}{\FOMLL}{\Deri'_1}{\sqns{\Gamma'}}{
      \vlhy{\quad}}{
      \vlhy{}}{
      \vlhy{}}}}
\end{equation*}
Note that $\fequ$ is required here since the permutation of formulas is implicit
in $\FOMLL$.

By transforming each sequent of $\Deri'_2$ into its corresponding formula, and by
considering the following rewriting rule:
\begin{equation*}
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, A}}{
    \vlhy{\sqns{\Gamma}}}}
\leadsto
\vlderivation{
  \vlin{\wrD}{}{\sqns{\form{\Gamma} \vlor A}}{
    \vlhy{\sqns{\form{\Gamma}}}}}
\end{equation*},
we obtain a derivation
\begin{equation*}
\vlderivation{
  \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
    \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\form{\Gamma'}}}{
      \vlhy{\quad}}{
      \vlhy{}}{
      \vlhy{}}}}
\end{equation*}
where $\Deri_1$ can be obtained from $\Deri'_1$
by applying the $\vlor$ rule.
\end{proof}

>>>>

\begin{proof}
  That $\graphof A$ is a logical cograph follows immediately from the definition and Theorem~\ref{thm:cograph}.
  The fact that every binder of $\graphof A$ is legal can be proved by
  structural induction on $A$.
  %% :
  %% \begin{itemize}
  %%   \item Trivial if $A$ is an atom.
  %%   \item The induction step uses the following fact:
  %%         if $M$ is a module of $\gG$ and $M' \subseteq M$, then $M'$ is a
  %%         module of $\gG$ if and only if it is a module of $\gG[X]$.
  %% \end{itemize}
  %% \todo{}\lutz{maybe mention also fn 2-5 in Dominic's arxiv-paper}
\end{proof}
%%\juihsuan{I'm still looking for a simple proof}
%
\begin{remark}
Note that $\graphof A$ is not necessarily a fograph if $A$ is not
rectified. If $A = (\forall x. p(x)) \vlor (\forall x. q(x))$, then
$\graphof A = \single x \ \single p(x) \ \single x \ \single q(x)$,
the scope of each $x$-binder contains all the vertices, in particular,
the other $x$-binder. On the other hand, there are non-rectified
formulas which are translated to fographs by $\graphof\cdot$. For
example, in the graph of $(\exists x. p(x)) \vlor (\exists x. q(x))$,
both $x$-binders are legal, as they are not in each other's
scope:
$\edgepairpic{x}{px}\mkern9mu\edgepairpic{x}{qx}$.
\end{remark}
%
%% For this reason, we call a formula
%% \bfit{clean} if it does not contain subformulas of the form $(\forall
%% x.A)\vlor(\forall x.B)$ or $(\exists x.A)\vlan(\exists x.B)$,
%% \todo{need the four cases as in clear definition in fopws? Consider
%% $(\forall x.A)\vlor(B\vlor (\forall x.C))$. But then Lemma 6 doesn't hold.
%% (fopws proves only that there's a surjection from clear formulas onto fographs,
%% whose kernel is the six formula equalities)} and no
%% $x$-quantified formula occurs as subformula of another $x$-quantified
%% formula. Then we have:
%
%% \begin{lemma}
%%   If $A$ is clean iff\/ $\graphof A$ is a fograph.
%% \end{lemma}
%
%% \begin{proof}
%%   Induction on $A$, using  Theorem~\ref{thm:cograph}.
%% \end{proof}
%
%% Even though for every formula $A$ we can obtain an
%% equivalent clean formula by simply renaming some bound variables, this
%% is not unique up to $\alpha$-conversion, as is the case for
%% rectified formulas.
%
\noindent We define a congruence relation $\fequ$ on formulas by the following equations:
\begin{equation}
  \label{eq:fequ}
  \begin{array}{c@{\hskip1.5em}c}
    A \vlan B \fequ B \vlan A
    &
    (A \vlan B) \vlan C \fequ A \vlan (B \vlan C) 
    \\
    A \vlor B \fequ B \vlor A
    &
    (A \vlor B) \vlor C \fequ A \vlor (B \vlor C)
    \\
    \forall x . \forall y . A \fequ \forall y . \forall x . A
    &
    \forall x . (A \vlor B) \fequ (\forall x . A) \vlor B
    \\
    \exists x . \exists y . A \fequ \exists y . \exists x . A
    &
    \exists x . (A \vlan B) \fequ (\exists x . A) \vlan B
  \end{array}
\end{equation}
where $x$ must not be free in $B$ in the last two equations.
Two formulas $A$ and $B$ are \bfit{equivalent} if $A \fequ B$. The following
theorem shows that the set of fographs can been seen as the quotient
$\FORM/\mathord\fequ$.
\begin{thm}
  Let $A,B$ be rectified formulas. Then
  $$
  A\fequ B \iff \graphof A =\graphof B
  $$
\end{thm}

@@

Let $\Pi$ be a $\FOMLL$ proof of a rectified sequent $\sqn\Gamma$ not containing $\fff$. We now show
how $\Pi$ is translated into a linked fograph
$\fographof\Pi=\tuple{\graphof\Gamma,\linkingof\Pi}$. We proceed
inductively, making a case analysis on the last rule in $\Pi$. At the
same time we are constructing a dualizer $\dsubstof\Pi$, so that in the
end we can conclude that $\fographof\Pi$ is in fact a fonet.
\begin{enumerate}
  \addtolength\itemsep{.7ex}
\item $\Pi$ is $\upsmash{\vlinf{\axr}{}{\sqn{a,\cneg a}}{}}$ : Then the only
  link is $\set{a, \dual{a}}$, and $\dsubstof\Pi$ is empty.
\item $\Pi$ is $\smash{\vlinf{\ttt}{}{\sqn{\ttt}}{}}$ : Then
  $\linkingof\Pi$ and $\dsubstof\Pi$ are both empty.
\item The last rule in $\Pi$ is $\vliinf{\mixr}{}{\sqn{\Gamma',\Gamma''}}{\sqn\Gamma'}{\sqn\Gamma''}$ :
  By induction hypothesis, we have proofs $\Pi'$ and $\Pi''$ of $\Gamma'$ and $\Gamma''$, respectively. We have
  $\graphof{\Gamma}=\graphof{\Gamma'}+\graphof{\Gamma''}$ and we can let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}\cup\linkingof{\Pi''}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}\cup\dsubstof{\Pi''}$.
\item The last rule in $\Pi$ is $\vlinf{\vlor}{}{\sqn{\Gamma_1,A\vlor
    B}}{\sqn{\Gamma_1,A,B}}$ : By induction hypothesis, there is
  proofs $\Pi'$ of $\Gamma'={\Gamma_1,A,B}$. We have
  $\graphof{\Gamma}=\graphof{\Gamma'}$ and let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}$.
\item The last rule in $\Pi$ is $\vliinf{\vlan}{}{\sqn{\Gamma_1,A\vlan
    B, \Gamma_2}}{\sqn{\Gamma_1,A}}{\sqn{B,\Gamma_2}}$ : By induction
  hypothesis, we have proofs $\Pi'$ and $\Pi''$ of
  $\Gamma'=\Gamma_1,A$ and $\Gamma''=B,\Gamma_2$, respectively. We
  have $\graphof{\Gamma}=\graphof{\Gamma_1}+\mbox{$(\graphof A\times\graphof
  B)$}+\graphof{\Gamma_2}$ and we let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}\cup\linkingof{\Pi''}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}\cup\dsubstof{\Pi''}$.
\item The last rule in $\Pi$ is
  $\vlinf{\exists}{}{\sqn{\Gamma_1,\exists x.A}}
  {\sqn{\Gamma_1,A\ssubst{x}{t}}}$ : By induction hypothesis, there is
  a $\Pi'$ of $\Gamma'={\Gamma_1,A\ssubst{x}{t}}$.  For each atom in
  $\Gamma'=\Gamma_1, A \ssubst{x}{t}$, there is a corresponding atom
  in $\Gamma=\Gamma_1, \exists x.A$. We can therefore define the
  linking $\linkingof{\Pi}$ from the linking $\linkingof{\Pi'}$ via
  this correspondence. Then, we let $\dsubstof\Pi$ be
  $\dsubstof{\Pi'}+\ssubst{x}{t}$. Since $\Gamma$ is rectified $x$
  does not yet occur in $\dsubstof{\Pi'}$. Hence $\dsubstof\Pi$ is a
  dualizer of $\fographof\Pi$.
\item The last rule in $\Pi$ is $\vlinf{\forall}{\text{($x$ not free
    in $\Gamma_1$)}}{\sqn{\Gamma_1,\forall x.A}}{\sqn{\Gamma_1,A}}$ :
  By induction hypothesis, there is a proof $\Pi'$ of
  $\Gamma'={\Gamma_1,A}$, which has the same atoms as in
  $\Gamma=\Gamma_1, \forall x.A$.  Hence, we can let
  $\linkingof\Pi\;=\;\linkingof{\Pi'}$ and
  $\dsubstof\Pi=\dsubstof{\Pi'}$. 
\end{enumerate}
%\end{itemize}

@@

\begin{lemma}\label{lem:skew->cw}
  Let $\gA$ and $\gB$ be fographs, let $\phi\colon\gA\to\gB$ be a skew
  bifibration, and let $A$ and $B$ be formulas with $\graphof A=\gA$
  and $\graphof B=\gB$. Then there are derivations
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wrD,\wfaD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{B}{
        \vlhy{A}}}
    \quand
    \vlderivation{
      \vlde{\set{\wrD,\wfaD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
        \vlhy{A\rsubstof\phi}}}
  \end{equation*}
  such that $\mapof{\rectif\Deri}=\phi$ and $\rectif{\Deri}$ is a
  rectification of $\Deri$, and $\rsubstof\phi$ is the substitution
  induced by $\phi$.
\end{lemma}

@@

We say a fograph homomorphism $\phi\colon\gG\to\gH$ is \bfit{full} if
for all $v,w\in\vG$, we have that $\phi(v)\phi(w)\in\eH$ implies
$vw\in\eG$.

\begin{lemma}\label{lem:fullinj->w}
  Let $\phi\colon\gG\to\gH$ be full and injective skew bifibration
  such that $\rsubstof\phi$ is the identity substitution, and let $G$
  and $H$ be formulas with $\graphof G=\gG$ and $\graphof H=\gH$. Then
  There is a derivation $\vlder{\set{\wrD,\wfaD,\fequ}}{}{H}{G}$.
\end{lemma}

\begin{proof}
   By \cite[Proposition~7.6.1]{str:07:RTA}, we have a
  derivation $\vlder{\set{\wrD,\PE\fequ}}{\DDeri}{\PE H}{\PE G}$.
  A subformula of $\PE B$ is called \emph{weak} if it has been
  introduced by a weakening. All subformulas of a weak formula are
  also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  \emph{weak pair} if $\PE B\fequ\Scons{B'\vlor B''}$ for some context
  $\Sconhole$. We can assume without loss of generality that
  whenever weak subformulas $B'$ and $B''$ form a weak pair, they have
  been introduced by the same instance of $\wrD$ in
  $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  brought into this form by simple rule permutations, and then
  collapsing several weakenings into a single one.}  Now we show that
  $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  delete all atoms in the scope of the corresponding quantifier,
  because $\phi'$ is a fibration on the binding graph. Hence, each line
  in $\PE\Deri_1$ is the propositional encoding $\PE P$ of a
  first-order formula $P$.  We now have to show
  that each instance of $\wrD$ is indeed a correct
  application in first-order logic. For this we have to inspect the
  cases a weakening happens inside a subformula $x\vlor C$ or $x\vlan
  C$ in $\PE\Deri_1$. There are the following cases:
  \begin{equation*}%\small
    \vlinf{}{}{S\cons{x\vlor D\vlor C}}{S\cons{x\vlor C}}
    \qquad
    \vlinf{}{}{S\cons{x\vlan (D\vlor C)}}{S\cons{x\vlan C}}
    \qquad
    \vlinf{}{}{S\cons{(x\vlor D)\vlan C}}{S\cons{x\vlan C}}
  \end{equation*}
  In the first case the weakening happens inside the scope of a
  $\forall$-quantifier, and in the second case inside the scope of a
  $\exists$-quantifier. Both cases are unproblematic in first-order
  logic. However, in the third case an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur.
  Thus we have a first order derivation
  $\vlder{\set{\wrD,\fequ}}{\Deri_1}{B}{B_1}$ with $\PE B_1=\PEp B_1$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:skew->cw}]
  By Lemma~\ref{lem:PE} we have $\gA=\graphof{\PE A}$ and
  $\gB=\graphof{\PE B}$.  Let $\vB'\subseteq\vB$ be the image of
  $\phi$, and let $\gB_1$ be the subgraph of $\gB$ induced by
  $\vB'$. Hence, we have two maps $\phi''\colon\gA\to\gB_1$ being a
  surjection and $\phi'\colon \gB_1\to\gB$ being an injection that
  reflects edges.  Both, $\phi'$ and $\phi''$ remain skew
  bifibrations. Let us first look at $\phi'$.  Let $\PEp B_1$ be the
  propositional formula obtained from $\PE B$ by removing all atoms
  that are not represented by vertices in $\vB'$. Then $\graphof{\PEp
    B_1}=\gB_1$.

@@

\begin{equation*}
\vlderivation{
  \vliin{\vlan}{}{\sqns{\Gamma, A \vlan B, \Delta}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A}}{
      \vlhy{\sqns{\Gamma}}}}{
    \vlhy{\sqns{B, \Delta}}}}
\leadsto
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, A \vlan B, \Delta}}{
    \vlin{\weakr}{}{\vlvdots}{
      \vlhy{\sqns{\Gamma}}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\vlor}{}{\sqns{\Gamma, A \vlor B}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A, B}}{
      \vlhy{\sqns{\Gamma, A}}}}}
\leadsto
\vlderivation{
  \vlin{\wrD}{}{\sqns{\Gamma, A \vlor B}}{
    \vlhy{\sqns{\Gamma, A}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\exists}{}{\sqns{\Gamma, \exists x.A}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A\ssubst{x}{t}}}{
      \vlhy{\sqns{\Gamma}}}}}
\leadsto
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, \exists x.A}}{
    \vlhy{\sqns{\Gamma}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\forall}{}{\sqns{\Gamma, \forall x.A}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A}}{
      \vlhy{\sqns{\Gamma}}}}}
\leadsto
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, \forall x.A}}{
    \vlhy{\sqns{\Gamma}}}}
\end{equation*}
\begin{equation*}
\vlderivation{
  \vlin{\conr}{}{\sqns{\Gamma, A}}{
    \vlin{\weakr}{}{\sqns{\Gamma, A, A}}{
      \vlhy{\sqns{\Gamma, A}}}}}
\leadsto
\vlderivation{
  \vlhy{\sqns{\Gamma, A}}}
\end{equation*}
Note that in the case of $\vlor$, we use the deep rule $\wrD$ which can be
permuted down over all the rules. By using these rewriting rules, we can eventually get
a derivation with all the instances of $\weakr$ and $\wrD$ at the bottom. Now observe
that the instances of $\conr$ in $\Deri$ can be transformed using the following
rule:

@@

Knowing that $\cD$ can be permuted down over all the rules of $\FOMLL$, we
eventually obtain a derivation:
\begin{equation*}
\vlderivation{
  \vlde{\set{\weakr, \wrD,\cD,\fequ}}{\Deri'_2}{\sqns{\Gamma}}{
    \vltrl{}{\FOMLL}{\Deri'_1}{\sqns{\Gamma_0}}{
      \vlhy{\quad}}{
      \vlhy{}}{
      \vlhy{}}}}
\end{equation*}
Note that $\fequ$ is required here since the permutation of formulas is implicit
in $\FOMLL$.

By transforming each sequent of $\Deri'_2$ into its corresponding formula, and by
considering the following rewriting rule:
\begin{equation*}
\vlderivation{
  \vlin{\weakr}{}{\sqns{\Gamma, A}}{
    \vlhy{\sqns{\Gamma}}}}
\leadsto
\vlderivation{
  \vlin{\wrD}{}{\sqns{\form{\Gamma} \vlor A}}{
    \vlhy{\sqns{\form{\Gamma}}}}}
\end{equation*},
we obtain a derivation
\begin{equation*}
\vlderivation{
  \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
    \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\Gamma'}}{
      \vlhy{\quad}}{
      \vlhy{}}{
      \vlhy{}}}}
\end{equation*}
where $\Gamma' = \form{\Gamma_0}$ and $\Deri_1$ can be obtained from $\Deri'_1$
by applying the $\vlor$ rule.
\end{proof}
